---
title: "Cougars_LA_County"
author: "Delores Chan"
date: "2024-02-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("rinat")
library(rinat)
library(tidyverse)
library(here)
library(janitor)
library(adehabitatHR)
library(dplyr)
library(ggmap)
```


```{r}
#la bounding box from: https://anthonylouisdagostino.com/bounding-boxes-for-all-us-counties/

bounds <- c(-118.951721, 32.75004, -117.646374,	34.823302)
cougars <- get_inat_obs(query="Puma concolor", place_id = 962, maxresults=5000)

plot(cougars$longitude, cougars$latitude)

#Error in get_inat_obs(query = "Puma concolor", bounds = bounds) : Your search returned zero results. Either your species of interest has no records or you entered an invalid search.
```

```{r}
# df with just id, lat, and long 

cougar_lat_long <- data.frame(cougars$id, cougars$observed_on, cougars$latitude, cougars$longitude)
```

```{r}
# df chnaing all id's to 1 (whole population)

la_county_cougars <- data.frame(cougars$longitude, cougars$latitude) %>% 
  add_column(id=1) %>% 
  rename(x=cougars.longitude, y=cougars.latitude)
```


```{r}
#coerce into spatial data frame 

cougar.sp <- la_county_cougars[!is.na(la_county_cougars$x) & !is.na(la_county_cougars$y),]
cougar.sp <- la_county_cougars[,c("id", "x", "y")] 

coordinates(cougar.sp) <- c("x", "y")
proj4string(cougar.sp) <- CRS("+proj=utm +zone=12 +datum=WGS84 +units=m +no_defs")
```

```{r}
# plotting with adehabitat

plot(cougar.sp, col = as.factor(cougar.sp@data$id), pch = 16)

# add a column of for the same number to appear as a whole population 
cougar.mcp <- mcp(cougar.sp, percent = 100)
cougar.mcp
```
```{r}
plot(cougar.sp, col=as.factor(cougar.sp@data$id), pch=16) #plot points
plot(cougar.mcp, col=alpha(1:5, 0.5), add=TRUE) #plot polygons
```

```{r}
#filter by year: looking at 2020 data, effects of pandemic? 

cougar_2020.sp <- cougars %>%  
  filter(quality_grade == 'research') %>% 
  dplyr::select(id, observed_on, latitude, longitude) %>% 
  mutate(id=1) %>% 
  filter(observed_on >='2020-01-01' & observed_on <= '2020-12-31')


coordinates(cougar_2020.sp) <- c("longitude", "latitude")
cougar_2020.mcp <- mcp(cougar_2020, percent = 100)
```


```{r}
plot(cougar_2020.sp, col=as.factor(cougar_2020.sp@data$id), pch=16) #plot points
plot(cougar_2020.mcp, col=alpha(1:5, 0.5), add=TRUE) #plot polygons
```
```{r}
## moratorium on rodenticide 


```


POPULATION level home range 
we are getting at population habitat USE 
selection is the proportion of use to availability 
build home range on subset (take 80/100) 
put the other 20 back and see where they landed 
break out by year? is the habitat use changing by 5 year periods? 
look at historical population data, compare crashes 

find meaningful discretization of time periods 

gganimate 

